The Power of Two Random Choices\\ \large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}

### 2014-02-09 article/article.tex
use \cdot for multiplication

A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy. Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.}~\cite{RS98}.

@@

Condsider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\lceil\frac{n}{d}\rceil$ \todo{simplify to n/d?}. We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins. In case a tie occurs, put the ball into the \emph{left-most} bin. That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with same minimal load, we put the ball into the bin with smallest index $i$. Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymetric tie-breaking} in combination with \emph{partitioning}\todo{check} reduces the bound for the gap to $\Theta \left( \frac{\log\log n}{d \log \Phi_d}\right)$ \todo{$\Phi_d$ is undefined}.

@@

The fullest bin contains
\begin{align}
\Theta \left( \frac{\log n}{\log \log n} \right)
\end{align} 
balls in the lightly loaded case w.h.p. Since the average load of the bins is 1, this is also a bound for the gap. 
%More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}. %\todo{really include Gonnet?}
In the \emph{heavily loaded case} the maximum load is 
\begin{align}
\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)
\end{align} 
w.h.p.~\cite{RS98}.

@@

\begin{theorem}
\label{theorem:agln}
If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
\begin{align}
\frac{\ln\ln n}{d \ln \Phi_d} \pm \Theta(1)
\end{align}
w.h.p.
\end{theorem}

\begin{theorem}
\label{theorem:algm}
Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most 
\begin{align}
\frac{\ln\ln n}{d \ln \Phi_d} + O(h)
\end{align}
w.h.p.
\end{theorem}

>>>>

A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy. Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right)$ with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.}~\cite{RS98}.

@@

Condsider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\lceil\frac{n}{d}\rceil$ \todo{simplify to n/d?}. We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins. In case a tie occurs, put the ball into the \emph{left-most} bin. That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with same minimal load, we put the ball into the bin with smallest index $i$. Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymetric tie-breaking} in combination with \emph{partitioning}\todo{check} reduces the bound for the gap to $\Theta \left( \frac{\log\log n}{d  \cdot \log \Phi_d}\right)$ \todo{$\Phi_d$ is undefined}.

@@

The fullest bin contains
\begin{align}
\Theta \left( \frac{\log n}{\log \log n} \right)
\end{align} 
balls in the lightly loaded case w.h.p. Since the average load of the bins is 1, this is also a bound for the gap. 
%More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}. %\todo{really include Gonnet?}
In the \emph{heavily loaded case} the maximum load is 
\begin{align}
\frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right)
\end{align} 
w.h.p.~\cite{RS98}.

@@

\begin{theorem}
\label{theorem:agln}
If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
\begin{align}
\frac{\ln\ln n}{d \cdot \ln \Phi_d} \pm \Theta(1)
\end{align}
w.h.p.
\end{theorem}

\begin{theorem}
\label{theorem:algm}
Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most 
\begin{align}
\frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(h)
\end{align}
w.h.p.
\end{theorem}
